---
title: "Processing of the data from: Altered oligodendrocyte heterogeneity in Multiple sclerosis revealed by single nuclei RNA sequencing Jaëkel, Agirre et al. 2019"
output: html_notebook
---


```{r}
library(scater)
library(Seurat)
library(velocyto.R)
```
load loom files fro velocyto
Load loom data
```{r}
# 
ldat_052 <- read.loom.matrices("10X_17_052_grch38.loom")
ldat_053 <- read.loom.matrices("10X_17_053_grch38.loom")
ldat_054 <- read.loom.matrices("10X_17_054_grch38.loom")
ldat_055 <- read.loom.matrices("10X_17_055_grch38.loom")
ldat_056 <- read.loom.matrices("10X_17_056_grch38.loom")
ldat_057 <- read.loom.matrices("10X_17_057_grch38.loom")
ldat_058 <- read.loom.matrices("10X_17_058_grch38.loom")
ldat_059 <- read.loom.matrices("10X_17_059_grch38.loom")
ldat_060 <- read.loom.matrices("10X_17_060_grch38.loom")
ldat_061 <- read.loom.matrices("10X_17_061_grch38.loom")
ldat_062 <- read.loom.matrices("10X_17_062_grch38.loom")
ldat_063 <- read.loom.matrices("10X_17_063_grch38.loom")
ldat_064 <- read.loom.matrices("10X_17_064_grch38.loom")
ldat_065 <- read.loom.matrices("10X_17_065_grch38.loom")
ldat_066 <- read.loom.matrices("10X_17_066_grch38.loom")
ldat_067 <- read.loom.matrices("10X_17_067_grch38.loom")
ldat_068 <- read.loom.matrices("10X_17_068_grch38.loom")
ldat_069 <- read.loom.matrices("10X_17_069_grch38.loom")
ldat_070 <- read.loom.matrices("10X_17_070_grch38.loom")
ldat_071 <- read.loom.matrices("10X_17_071_grch38.loom")



nuclei_mat <- as.matrix(cbind(ldat_052$spliced , ldat_053$spliced , ldat_054$spliced , ldat_055$spliced , ldat_056$spliced , ldat_057$spliced , ldat_058$spliced , ldat_059$spliced , ldat_060$spliced , ldat_061$spliced , ldat_062$spliced , ldat_063$spliced , ldat_064$spliced , ldat_065$spliced , ldat_066$spliced , ldat_067$spliced , ldat_068$spliced , ldat_069$spliced , ldat_070$spliced , ldat_071$spliced ) )
                        

save( nuclei_mat , file = "Nuclei_spliced_roche.Robj")


nuclei_unspliced_mat <- as.matrix(cbind(ldat_052$unspliced , ldat_053$unspliced , ldat_054$unspliced , ldat_055$unspliced , ldat_056$unspliced , ldat_057$unspliced , ldat_058$unspliced , ldat_059$unspliced , ldat_060$unspliced , ldat_061$unspliced , ldat_062$unspliced , ldat_063$unspliced , ldat_064$unspliced , ldat_065$unspliced , ldat_066$unspliced , ldat_067$unspliced , ldat_068$unspliced , ldat_069$unspliced , ldat_070$unspliced , ldat_071$unspliced  ))  

save( nuclei_unspliced_mat , file =  "Nuclei_unspliced_roche.Robj")


```


load the intronexon matrix
```{r }

load("Nuclei_spliced_roche.Robj")
load("Nuclei_unspliced_roche.Robj")


rownames(nuclei_mat) <- make.names( rownames(nuclei_mat) , unique=T )
rownames(nuclei_unspliced_mat) <- make.names( rownames(nuclei_unspliced_mat) , unique=T )

matlist <- list(nuclei_mat, nuclei_unspliced_mat)
nuclei_combmat <- matlist[[1]]+matlist[[2]]



save( nuclei_combmat , file =  "Nuclei_intronexon_roche.Robj")


```
create an annotation file form Cellranger output info: filtered barcodes fro meach sample from cellranger
```{bash }


 awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_053_roche_grch38:"a[1]"x\tCO28\tCO28\tCtrl\tCtrl\tCO28" }}'  barcodes_052_roche_filtered_samples.tsv > barcodes_052_roche_filtered_samples_3.tsv
 
awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_054_roche_grch38:"a[1]"x\tCO25\tCO25\tCtrl\tCtrl\tCO25" }}'  barcodes_053_roche_filtered_samples.tsv > barcodes_053_roche_filtered_samples_3.tsv 
 
awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_055_roche_grch38:"a[1]"x\tMS176_NAWM\tMS176_NAWM\tMS\tNAWM\tMS176" }}'  barcodes_054_roche_filtered_samples.tsv > barcodes_054_roche_filtered_samples_3.tsv 
 
awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_056_roche_grch38:"a[1]"x\tMS176_RM\tMS176_RM\tMS\tRM\tMS176" }}'  barcodes_055_roche_filtered_samples.tsv > barcodes_055_roche_filtered_samples_3.tsv 

awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_056_roche_grch38_4:"a[1]"x\tMS176_CI\tMS176_CI\tMS\tCI\tMS176" }}'  barcodes_056_roche_filtered_samples.tsv > barcodes_056_roche_filtered_samples_3.tsv 
 

 
awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_057_roche_grch38:"a[1]"x\tMS176_CA\tMS176_CA\tMS\tCA\tMS176" }}'  barcodes_057_roche_filtered_samples.tsv > barcodes_057_roche_filtered_samples_3.tsv 


awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_058_roche_grch38:"a[1]"x\tCO39\tCO39\tCtrl\tCtrl\tCO39" }}'  barcodes_058_roche_filtered_samples.tsv > barcodes_058_roche_filtered_samples_3.tsv 
 
awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_059_roche_grch38:"a[1]"x\tCO14\tCO14\tCtrl\tCtrl\tCO14" }}'  barcodes_059_roche_filtered_samples.tsv > barcodes_059_roche_filtered_samples_3.tsv 
 
awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_060_roche_grch38:"a[1]"x\tMS242_CA5\tMS242_CA5\tMS\tCA\tMS242" }}'  barcodes_060_roche_filtered_samples.tsv > barcodes_060_roche_filtered_samples_3.tsv 
 
awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_061_roche_grch38:"a[1]"x\tMS242_CI\tMS242_CI\tMS\tCI\tMS242" }}'  barcodes_061_roche_filtered_samples.tsv > barcodes_061_roche_filtered_samples_3.tsv 


awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_062_roche_grch38:"a[1]"x\tMS242_CA2\tMS242_CA2\tMS\tCA\tMS242" }}'  barcodes_062_roche_filtered_samples.tsv > barcodes_062_roche_filtered_samples_3.tsv 

awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_063_roche_grch38:"a[1]"x\tMS242_RM\tMS242_RM\tMS\tRM\tMS242" }}'  barcodes_063_roche_filtered_samples.tsv > barcodes_063_roche_filtered_samples_3.tsv 

awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_064_roche_grch38:"a[1]"x\tSD48/16\tSD48/16\tCtrl\tCtrl\tSD48/16" }}'  barcodes_064_roche_filtered_samples.tsv > barcodes_064_roche_filtered_samples_3.tsv 


awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_065_roche_grch38:"a[1]"x\tMS122_NAWM\tMS122_NAWM\tMS\tNAWM\tMS122" }}'  barcodes_065_roche_filtered_samples.tsv > barcodes_065_roche_filtered_samples_3.tsv 

awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_066_roche_grch38:"a[1]"x\tMS122_CI\tMS122_CI\tMS\tCI\tMS122" }}'  barcodes_066_roche_filtered_samples.tsv > barcodes_066_roche_filtered_samples_3.tsv 

awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_067_roche_grch38:"a[1]"x\tMS122_A\tMS122_A\tMS\tA\tMS122" }}'  barcodes_067_roche_filtered_samples.tsv > barcodes_067_roche_filtered_samples_3.tsv 


awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_068_roche_grch38:"a[1]"x\tMS121_NAWM\tMS121_NAWM\tMS\tNAWM\tMS121" }}'  barcodes_068_roche_filtered_samples.tsv > barcodes_068_roche_filtered_samples_3.tsv 


awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_069_roche_grch38:"a[1]"x\tMS121_A3\tMS121_A3\tMS\tA\tMS121" }}'  barcodes_069_roche_filtered_samples.tsv > barcodes_069_roche_filtered_samples_3.tsv 


awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_070_roche_grch38:"a[1]"x\tMS121_A2\tMS121_A2\tMS\tA\tMS121" }}'  barcodes_070_roche_filtered_samples.tsv > barcodes_070_roche_filtered_samples_3.tsv 

awk '{ split($1,a,"-") ; if ( a[2] == 1 ) { print "10X_17_071_roche_grch38:"a[1]"x\tMS121_CA\tMS121_CA\tMS\tCA\tMS121" }}'  barcodes_071_roche_filtered_samples.tsv > barcodes_071_roche_filtered_samples_3.tsv 

########################################################################
cat *3.tsv  > 10X_Nuclei_roche_networkAnnotableFile_corrected.txt

```
QC and processing of the UMI matrix
```{r }
networkAnnotableFile <- read.table("10X_Nuclei_roche_networkAnnotableFile_corrected.txt")

header <- c( "cell" , "original" , "sample" , "condition" , "lesion" , "patient" )
colnames(networkAnnotableFile) <- header

table(grepl("MALAT1",rownames(nuclei_combmat)))
nuclei_combmat <- nuclei_combmat[! grepl("MALAT1",rownames(nuclei_combmat)),]

table(grepl("XIST",rownames(nuclei_combmat)))
nuclei_combmat <- nuclei_combmat[! grepl("XIST",rownames(nuclei_combmat)),]


sceset_intronexon <- SingleCellExperiment(
    assays = list(counts = nuclei_combmat ), 
    colData = networkAnnotableFile
)



keep_feature <- rowSums(counts(sceset_intronexon) > 0) > 0
sceset_intronexon <- sceset_intronexon[keep_feature,]

is.mito <- grepl("^MT.", rownames(sceset_intronexon))
summary(is.mito)
#there is  mitochondrial
sceset_intronexon <- calculateQCMetrics(sceset_intronexon , feature_controls=list( Mt=is.mito)  )

ave.counts <- calcAverage(sceset_intronexon)
hist(log10(ave.counts), breaks=100, main="", col="grey80", 
    xlab=expression(Log[10]~"average count"))


rowData(sceset_intronexon)$ave.count <- ave.counts
to.keep <- ave.counts > 0
sceset_intronexon <- sceset_intronexon[to.keep,]
summary(to.keep)



save( sceset_intronexon , file =  "Nuclei_intronexon_sceset.Robj")


```
# Remove cells with total transcript count below 500
```{r }
summary(sceset_intronexon$total_counts)

hist(sceset_intronexon$total_counts, 100)
abline(v = 300, col = "red", lwd = 2)

summary(sceset_intronexon$total_features)

hist(sceset_intronexon$total_features, 100)
abline(v = 200, col = "red", lwd = 2)


sceset_intronexon <- sceset_intronexon[, sceset_intronexon$total_counts > 300]
sceset_intronexon <- sceset_intronexon[, sceset_intronexon$total_features > 200]

dim(sceset_intronexon)
```
## Exclude genes that don't have a count above 1 in at least 2 cells
```{r }
filter_genes <- apply(counts(sceset_intronexon), 1, function(x) length(x[x >= 1]) >= 2)
sceset_intronexon <- sceset_intronexon[filter_genes, ]
dim(sceset_intronexon)

save( sceset_intronexon , file =  "Nuclei_intronexon_sceset_filtered.Robj")

```
Create two objects: one where all the samples are gonna be trated separatelly and a common one for sup.fig5

All samples together:
```{r}
sces_together <- as.matrix(counts(sceset_intronexon))

nuclei_together_seurat <- CreateSeuratObject(raw.data = sces_together , 
                          min.cells = 3, min.genes = 200, project = "MSCtrl")
 


mito.genes <- grep(pattern = "^MT.", x = rownames(x = nuclei_together_seurat@data), value = TRUE)

percent.mito <- Matrix::colSums(nuclei_together_seurat@raw.data[mito.genes, ])/Matrix::colSums(nuclei_together_seurat@raw.data)

nuclei_together_seurat <- AddMetaData(object = nuclei_together_seurat, metadata = percent.mito, col.name = "percent.mito")


annotation_roche <- read.table("10X_Nuclei_roche_networkAnnotableFile_corrected_header.txt", stringsAsFactors=F)

nuclei_together_seurat <- AddMetaData(object = nuclei_together_seurat , metadata = annotation_roche, col.name = "original")  
nuclei_together_seurat <- AddMetaData(object = nuclei_together_seurat , metadata = annotation_roche, col.name = "sample")  
nuclei_together_seurat <- AddMetaData(object = nuclei_together_seurat , metadata = annotation_roche , col.name = "condition")  
nuclei_together_seurat <- AddMetaData( object = nuclei_together_seurat , metadata = annotation_roche , col.name = "lesion" )
nuclei_together_seurat <- AddMetaData( object = nuclei_together_seurat , metadata = annotation_roche , col.name = "patient" )



Nuclei_intexon_seurat_filter <- FilterCells(object = nuclei_together_seurat, subset.names = c("nGene" , "percent.mito"),
    low.thresholds = c(200, -Inf), high.thresholds = c(3500, 0.20))

#write.table( Nuclei_intexon_seurat@meta.data , file =  "/media/eneritz/Elements1/Nuclei_MS/uppmax_calc/Nuclei_intronexon_seurat_meta-data.txt", sep="\t",row.names = T, col.names = TRUE, quote = FALSE)

 Nuclei_intexon_seurat_norm<-NormalizeData(object = Nuclei_intexon_seurat_filter , normalization.method = "LogNormalize", scale.factor = 10000)

 Nuclei_intexon_seurat_norm <- ScaleData( Nuclei_intexon_seurat_norm, vars.to.regress =c("nUMI" , "percent.mito"  ), display.progress = FALSE)
 
Nuclei_intexon_seurat_norm <- FindVariableGenes(object =Nuclei_intexon_seurat_norm , do.plot = F )



nuclei_together_seurat  <- Nuclei_intexon_seurat_norm

nuclei_together_seurat <- SetAllIdent(object = nuclei_together_seurat , id = "patient")
VlnPlot(object = nuclei_together_seurat , features.plot = c("nGene", "nUMI", "percent.mito"), nCol = 1 , point.size.use = 0.05 ) 
 
nuclei_together_seurat <- SetAllIdent(object = nuclei_together_seurat , id = "lesion")
VlnPlot(object = nuclei_together_seurat , features.plot = c("nGene", "nUMI", "percent.mito"), nCol = 1 , point.size.use = 0.5 ) 


par(mfrow = c(1, 2))
GenePlot(object = nuclei_together_seurat , gene1 = "nUMI", gene2 = "percent.mito")
GenePlot(object = nuclei_together_seurat , gene1 = "nUMI", gene2 = "nGene")

nuclei_together_seurat <- SetAllIdent(object = nuclei_together_seurat , id = "sample")


VlnPlot(object = nuclei_together_seurat , features.plot = c("nGene", "nUMI", "percent.mito"), nCol = 3 , point.size.use = 1) 
 
par(mfrow = c(1, 2))
GenePlot(object = nuclei_together_seurat , gene1 = "nUMI", gene2 = "percent.mito")
GenePlot(object = nuclei_together_seurat , gene1 = "nUMI", gene2 = "nGene")

save( nuclei_together_seurat , file =  "Nuclei_togetherQC_intronexon_sceset_roche_191118.Robj")
###rebuttal figures:
Roche_seq_QC_vioplot_patient.eps
Roche_seq_QC_vioplot_patient_2.eps
VlnPlot(object = nuclei_together_seurat , features.plot = c("nGene", "nUMI"), nCol = 1 , point.size.use = 0.02 ) 
VlnPlot(object = nuclei_together_seurat , features.plot = c("nGene", "nUMI"), nCol = 1 , point.size.use = 0 ) 
Roche_seq_QC_vioplot_patient.eps
Roche_seq_QC_vioplot_patient_2.eps

VlnPlot(object = nuclei_together_seurat , features.plot = c("nGene", "nUMI"), nCol = 1 , point.size.use = 0.02 , x.lab.rot =  T) 

VlnPlot(object = nuclei_together_seurat , features.plot = c("nGene", "nUMI"), nCol = 1 , point.size.use = 0 , x.lab.rot =  T) 

Roche_seq_QC_vioplot_sample.eps
Roche_seq_QC_vioplot_sample_2.eps
#1210x744
```
PCA clustering supplementary figure7 : all samples together without CCA
```{r}
#PCA AND CLUSTERING WITHOUT REGRESS OUT

Nuclei_intexon_seurat_normAll <- RunPCA(object = nuclei_together_seurat , pc.genes = nuclei_together_seurat@var.genes, pcs.compute = 20, pcs.print = 1:10, maxit = 500, weight.by.var = FALSE)

PrintPCA(object = Nuclei_intexon_seurat_normAll, pcs.print = 1:10, genes.print = 5, use.full = FALSE)

VizPCA(object = Nuclei_intexon_seurat_normAll, pcs.use = 1:10)


PCHeatmap(object = Nuclei_intexon_seurat_normAll, pc.use = 1:15, cells.use = 500, do.balanced = TRUE, 
    label.columns = FALSE, use.full = FALSE)

PCElbowPlot(object = Nuclei_intexon_seurat_normAll)

Nuclei_intexon_seurat_normAll  <- FindClusters(object = Nuclei_intexon_seurat_normAll, reduction.type = "pca", dims.use = 1:20, 
			  resolution = 4 , print.output = 0, save.SNN = TRUE , force.recalc = TRUE )

Nuclei_intexon_seurat_normAll  <- FindClusters(object = Nuclei_intexon_seurat_normAll, reduction.type = "pca", dims.use = 1:20, 
			  resolution = 2 , print.output = 0, save.SNN = TRUE , force.recalc = TRUE )

Nuclei_intexon_seurat_normAll  <- FindClusters(object = Nuclei_intexon_seurat_normAll, reduction.type = "pca", dims.use = 1:20, 
			  resolution = 0.8 , print.output = 0, save.SNN = TRUE , force.recalc = TRUE )

save( Nuclei_intexon_seurat_normAll , file =  "/proj/uppstore2017150/private/eneritz/roche18/Nuclei_intronexon_sceset_roche_PCA_191118.Robj")

```

dimension reduction
```{r}
Nuclei_intexon_seurat_normAll <- RunTSNE(object = Nuclei_intexon_seurat_normAll , dims.use = 1:20, do.fast = TRUE)
```
CLUSTERING in different resolutions : supplementary figure7 
```{r}

Nuclei_intexon_seurat_normAll <- SetAllIdent(object = Nuclei_intexon_seurat_normAll , id = "res.0.8")
TSNEPlot(object = Nuclei_intexon_seurat_normAll , do.label = TRUE, pt.size = 0.2 )

Nuclei_intexon_seurat_normAll <- SetAllIdent(object = Nuclei_intexon_seurat_normAll , id = "res.4")
TSNEPlot(object = Nuclei_intexon_seurat_normAll , do.label = TRUE, pt.size = 0.2 )

Nuclei_intexon_seurat_normAll <- SetAllIdent(object = Nuclei_intexon_seurat_normAll , id = "res.2")
TSNEPlot(object = Nuclei_intexon_seurat_normAll , do.label = TRUE, pt.size = 0.2 )
Nuclei_intexon_seurat_normAll <- SetAllIdent(object = Nuclei_intexon_seurat_normAll , id = "sample")
TSNEPlot(object = Nuclei_intexon_seurat_normAll , do.label = TRUE, pt.size = 0.2)

Nuclei_intexon_seurat_normAll <- SetAllIdent(object = Nuclei_intexon_seurat_normAll , id = "patient")
TSNEPlot(object = Nuclei_intexon_seurat_normAll , do.label = F, pt.size = 0.2 )

Nuclei_intexon_seurat_normAll <- SetAllIdent(object = Nuclei_intexon_seurat_normAll , id = "lesion")
TSNEPlot(object = Nuclei_intexon_seurat_normAll , do.label = F, pt.size = 0.2 )

Nuclei_intexon_seurat_normAll <- SetAllIdent(object = Nuclei_intexon_seurat_normAll , id = "condition")
TSNEPlot(object = Nuclei_intexon_seurat_normAll , do.label = F, pt.size = 0.2 )

save( Nuclei_intexon_seurat_normAll , file =  "Nuclei_intronexon_sceset_roche_PCA_191118.Robj")


```
Processing of the final expression and annotations matrix:
###########################################################################################################3
create a seurat object for each of the individual 20 samples: Trying to rmeove osme possible batch effects
```{r}

load("Nuclei_intronexon_sceset_filtered.Robj")


sceset_intronexon <- lapply(unique(sceset_intronexon$sample), function(i) {
  sceset_intronexon[, sceset_intronexon$sample == i]
})


sces2 <- lapply(sceset_intronexon, function(w) {
  tmp <- as.matrix(counts(w))
  colnames(tmp) <- paste0(colData(w)$sample, "__" , 
                          colData(w)$cell)
                  
  tmp
})

```

seurat object for each sample

```{bash}

awk '{ print $2"__"$1"\t"$2"\t"$3"\t"$4"\t"$5"\t"$6 } '  10X_Nuclei_roche_networkAnnotableFile_corrected_header.txt > 10X_Nuclei_roche_networkAnnotableFile_header.txt
```

```{r}

annotation_roche <- read.table("10X_Nuclei_roche_networkAnnotableFile_header.txt", stringsAsFactors=F)

## Create Seurat object for each sample
seurats2 <- lapply(sces2, function(w) {
  s <- CreateSeuratObject(raw.data = w, 
                          min.cells = 3, min.genes = 200, project = "MSCtrl")
 
   mito.genes <- grep(pattern = "^MT.", x = rownames(x = s@data), value = TRUE)
 
     percent.mito <- Matrix::colSums(s@raw.data[mito.genes, ])/Matrix::colSums(s@raw.data)
s <- AddMetaData(object = s , metadata = percent.mito, col.name = "percent.mito")  
     s <- AddMetaData(object = s , metadata = annotation_roche, col.name = "original")  
s <- AddMetaData(object = s , metadata = annotation_roche, col.name = "sample")  
s <- AddMetaData(object = s , metadata = annotation_roche , col.name = "condition")  
s <- AddMetaData( object = s , metadata = annotation_roche , col.name = "lesion" )
s <- AddMetaData( object = s , metadata = annotation_roche , col.name = "patient" )

s <- FilterCells(object = s, subset.names = c("nGene" , "percent.mito"), 
    low.thresholds = c(200, -Inf), high.thresholds = c(6000, 0.20)) ##maybe 3500

  s <- NormalizeData(s)
  
  s <- ScaleData(s, vars.to.regress =c("nUMI" , "percent.mito" ), display.progress = FALSE)
 
  s@meta.data$orig.ident <- s@meta.data$sample
  FindVariableGenes(s, do.plot = FALSE)
})


#############################################
save( seurats2 , file =  "Nuclei_intronexon_sceset_roche_filtered_seurats2.Robj")

```
run CCA using the grouping from the samples
```{r}
seurats <- seurats2

## Find highly variable genes as the union of the top HVGs from each sample 
hvgs_sample <- unique(unlist(lapply(seurats, function(w) rownames(head(w@hvg.info, n = 1000)))))

table(grepl("^RP",(hvgs_sample)))

hvgs_sample <- hvgs_sample[! grepl("^RP",(hvgs_sample))]
## Remove batch effects and merge samples
seurat_sample <- NULL
discarded_cells_sample <- c()
Ns <- seq_len(length(seurats))
while(length(Ns) > 0) {
  if (is.null(seurat_sample)) {
    seurat_sample <- RunCCA(object = seurats[[Ns[1]]], object2 = seurats[[Ns[2]]], genes.use = hvgs_sample)
    Ns <- Ns[-(1:2)]
  } else {
    seurat_sample <- RunCCA(object = seurat_sample, object2 = seurats[[Ns[1]]], genes.use = hvgs_sample)
    Ns <- Ns[-1]
  }
  seurat_sample <- CalcVarExpRatio(seurat_sample, reduction.type = "pca", 
                            grouping.var = "sample", dims.use= 1:15)
  seuratdiscarded_sample <- SubsetData(seurat_sample, subset.name = "var.ratio.pca", accept.high = 0.5)
  print(table(seuratdiscarded_sample@meta.data$sample))
  discarded_cells_sample <- c(discarded_cells_sample, rownames(seuratdiscarded_sample@meta.data))
  seurat_sample <- SubsetData(seurat_sample, subset.name = "var.ratio.pca", accept.low = 0.5)
  seurat_sample <- AlignSubspace(seurat_sample, reduction.type = "cca", 
                          grouping.var = "sample", dims.align = 1:15)
  seurat_sample@meta.data$sample <- paste(unique(seurat_sample@meta.data$sample), collapse = "_")
}

#######################################################################################################
save( seurat_sample , file =  "Nuclei_intronexon_sceset_filtered_roche_CCA.Robj")

save( hvgs_sample , file =  "HVGS1K_Nuclei_intronexon_sceset_filtered_roche_CCA.Robj")


```
findclusters for grouped samples sample CCA
```{r}
Nuclei_intex_cca_sample <- seurat_sample

## Run t-SNE on the merged data and find clusters
Nuclei_intex_cca_sample <- RunPCA(object = Nuclei_intex_cca_sample , pc.genes = hvgs_sample , do.print = TRUE, pcs.print = 1:25, 
    genes.print = 5)
VizPCA(object = Nuclei_intex_cca_sample, pcs.use = 1:2)
PCElbowPlot(object = Nuclei_intex_cca_sample )

#PCHeatmap(object = Nuclei_intex_cca_sample , pc.use = 1:25, cells.use = 50, do.balanced = TRUE, 
#    label.columns = FALSE, use.full = FALSE)
```
#clustering by sample CCA at different resolutions 
```{r}


Nuclei_intex_cca_sample <- RunTSNE(object = Nuclei_intex_cca_sample , reduction.use = "cca.aligned", dims.use = 1:15,    do.fast = TRUE)


Nuclei_intex_cca_sample <- FindClusters(object = Nuclei_intex_cca_sample , reduction.type = "cca.aligned", dims.use = 1:15, 
                       save.SNN = TRUE, print.output = FALSE , resolution = 2 )

Nuclei_intex_cca_sample <- FindClusters(object = Nuclei_intex_cca_sample , reduction.type = "cca.aligned", dims.use = 1:15, 
                       save.SNN = TRUE, print.output = FALSE , resolution = 4 )

Nuclei_intex_cca_sample <- FindClusters(object = Nuclei_intex_cca_sample , reduction.type = "cca.aligned", dims.use = 1:15, 
                       save.SNN = TRUE, print.output = FALSE , resolution = 0.8 , force.recalc = T )





save( Nuclei_intex_cca_sample , file =  "Nuclei_intronexon_filtered_CCA_sample_clusters.Robj")

```
Final annotations can be find in the github repository/EGA accession or the https://ki.se/en/mbb/oligointernode webpage, the celltypes were asigned manually and confirmed based on previous knowledge of markers and finally validated with basescope. For any doubt contact through the issues or eneritz.agirre@ki.se

